---
title: "Indoor Positioning Raw Data Cleaning"
author: "STAT4/510 Wasabees Group"
date: "2023-11-30"
output: pdf_document
---

1. Describe your data
2. Describe the basic variable component of your data
3. Report the various findings you have established so far, with interprtation (include discussion on what you find and add how useful it is to your project objective)
4. Discuss any challenges you encountered, and ways by which you handled these.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

1. Introduction

Indoor position systems (IPS) development is an active area of research that can be used in numerous settings. Part of efforts to develop, calibration and model these systems are achieved throught the use of WIFI signals. The following report, describes and characterizes a large dataset compiled in a 15 by 36 meter area that contains six (wifi routers) access points.  


```{r}

#Read the offline data txt file, with each row in the txt file becoming a text string
raw_offline = readLines("Data/offline.final.trace.txt") # Replace with the actual path in your directory
str(raw_offline) 

```


```{r clean_#_rows}

#The txt file contains multiple rows starting with # (for example, rows 1, 2, 3) which does not share the same structure as other rows. The code below gets rid of these rows.

clean_offline <- raw_offline[substr(raw_offline,start=1, stop=1) != "#"]
str(clean_offline)

#After this code, clean_offline has 146,080 rows, less than the original 151,392 rows


```


```{r split_line}
#In clean_offline, each row now has the same structure:
#Semicolons separate between variables
#Equal signs separate variable names and values
#For the signal readings, readings of the same access point are separated by commas

#The code below took all the separation out of one line (line 1)
line = strsplit(clean_offline[1], "[; , =]") [[1]]
line

#The code below takes the time, id, pos, and degree readings out of "line" and arrange them in a matrix
line[c(2,4,6,7,8,10)] %>% 
  matrix(ncol=6,byrow=TRUE) %>% view()

#The code below takes the mac, signal, frequency, and device type reading out of "line" and arrange them in a matrix
line[-(1:10)] %>%
  matrix(ncol=4, byrow=TRUE) %>% view()

#After this, we can combine the two matrices to create one row of data for our ideal data frame. 

#The code below attempts to put all the above operations in this code chunk in one function
clean_function = function(x) {
  line = strsplit(x, "[; , =]") [[1]]
  a = line[-(1:10)] %>% matrix(ncol=4, byrow=TRUE)  
  b = line[c(2,4,6,7,8,10)] %>% matrix(ncol=6, nrow=nrow(a), byrow=TRUE) 
  cbind(b,a)
}

#Test clean_function with row 1 in clean_offline
lapply(clean_offline[1], clean_function) [[1]] %>% view()

```

```{r}
#This code chunk will use clean_function for all rows in clean_offline and rbind them together
```


# Plot the orientation of the measurement device
plot(ecdf(offline$orientation))
